Nitric oxide feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-05-29 Manuscript/Manuscript.Rmd
MS: modelling section edit

>>>>

To further analyse the dynamics of responses to UV light and formally describe cPRC phototransduction, synaptic connections and NO retrograde signalling, we developed a mixed cellular-circuit-level mathematical model. We used our Ca2+ imagining data of cPRC, INNOS and INRGW cells collected in wild type, *NOS* knockout, and *NIT-GC2* morphant larvae to formulate assumptions that are the basis of the proposed model.

We model a direct c-opsin1-dependent response to UV leading to the initial rise in Ca2+ levels. The mechanism of this signal is not known but may include Gα and Gβ signalling, CNG channel activation or other pathways. We further assume that in cPRC cells, Ca2+ levels increase with cGMP. We model a decrease in cGMP and Ca2+ levels due to a NIT-GC2-dependent pathway. The effect of NO in cPRC cells is captured by another term that describes a NIT-GC1 and NO-dependent increase in cGMP. Based on the synaptic connectome, we infer a feedforward coupling between cPRC cells and INNOS and INRGW cells. The proposed synaptic coupling is UV-dependent, decays linearly and is suppressed in NIT-GC2 morphants.
